Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-16 nips_2018.tex
rename fold to foldr

\begin{figure*}[h]
  \centering
  \begin{tabular}{rc}\toprule
    \sc{Task}& \begin{tabular}{rcl}
      Nancy FreeHafer&$\longrightarrow$&Dr. Nancy\\
      Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
      Pushmeet Kohli&$\longrightarrow$&???
    \end{tabular}
      \\\midrule
    \sc{Program}& $f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}
      \\\midrule
    \sc{DSL}& \begin{tabular}{l}
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y)))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
      \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
      \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}
      \\\bottomrule
  \end{tabular}\right.
\end{figure*}

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we initially provide a generic
 set of programming primitives in a Lisp-like language.
% including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Lists& \code{(fold nil (lambda (a b) (cons a b)))}\\
   &\hspace{0.5cm} \emph{(reverses a list)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
    \vspace{-4em}
 \end{wrapfigure}

@@

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\vspace{-0.7cm}\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
%% Although this  may seem 
%% very similar to estimating the parameters of a probabilistic context free grammar,
%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
%% our DSLs are context-sensitive due to the presence of variables
%% in the programs and also due to the polymorphic typing system.
%% In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
\begin{figure*}
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed subexpression\\\midrule
    \begin{tabular}{l}
      \code{(lambda (a b) (fold b (cons "," a)}\\
      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
      \code{(lambda (a b) (fold a b (lambda (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(fold a b (lambda (x z) }\\
      \phantom{\code{(fold a b }}\code{(cons x z)))}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}). Here, the proposed subexpression appends two lists.} 
  \label{fragmentExample}
\end{figure*}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to list processing (Section~\ref{textSection}) and text editing (Section~\ref{listSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
and initially provide the system with a generic
set of list processing primitives: \code{fold}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

@@

A complete repertoire of higher-order functions is a staple of functional programming standard libraries.
Although we provided \system with some of the standard higher-order functions,
like \code{fold} and \code{unfold},
we did not include others.
When trained on these list functions,
our system rediscovers
and then reuses the higher-order function \code{filter} \textbf{Lucas: not sure if it actually does this! It would be cool if we can write something like this though.}

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %(lambda (lambda (fold $0 $0 (lambda (lambda (if (char-eq? $1 $3) empty (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some string editing learned subroutines }\label{textPrimitives}
  \end{table}

>>>>

\begin{figure*}[h]
  \centering
  \begin{tabular}{rc}\toprule
    \sc{Task}& \begin{tabular}{rcl}
      Nancy FreeHafer&$\longrightarrow$&Dr. Nancy\\
      Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
      Pushmeet Kohli&$\longrightarrow$&???
    \end{tabular}
      \\\midrule
    \sc{Program}& $f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}
      \\\midrule
    \sc{DSL}& \begin{tabular}{l}
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y)))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a)}\\
      \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
      \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}
      \\\bottomrule
  \end{tabular}\right.
\end{figure*}

@@

We apply \system to three domains:
 symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing; and Lisp-style functions on lists.
 For each of these we initially provide a generic
 set of programming primitives in a Lisp-like language.
% including conditionals, variables, and higher-order functions.
 Our algorithm then discovers
 its own domain-specific vocabulary for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{wrapfigure}{r}{0.6\textwidth}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   %% Circuits&\code{(nand (nand a a) (nand b b))}\\
   %% &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(map (lambda (x) (if (= x a) b x)))}\\
   &\hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
   Lists& \code{(foldr nil (lambda (a b) (cons a b)))}\\
   &\hspace{0.5cm} \emph{(reverses a list)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
    \vspace{-4em}
 \end{wrapfigure}

@@

        \STATE $\mathcal{D},\theta\gets $induceDSL$(\{\mathcal{F}^{\theta}_x\cup\mathcal{F}^{q}_x\}_{x\in X})$  \hspace{1.1cm} \footnotesize{\textbf{(Compress)}}
        %% \STATE Define $Q_x(z) \propto \begin{cases}
        %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
        %%   0&x\not \in \mathcal{F}_x
        %% \end{cases}
        \ENDFOR
        \STATE \textbf{return} $\mathcal{D},\theta,q$
      \end{algorithmic}
    \end{algorithm}
  \end{minipage}
\vspace{-0.7cm}\end{wrapfigure}%% To appropriately score each proposed $\mathcal{D}$ we must reestimate the weight vector $\theta$.
%% Although this  may seem 
%% very similar to estimating the parameters of a probabilistic context free grammar,
%% for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
%% our DSLs are context-sensitive due to the presence of variables
%% in the programs and also due to the polymorphic typing system.
%% In the Supplement we derive a tractable MAP estimator for $\theta$.
Putting all these ingredients together, Alg.~\ref{mainAlgorithm} describes how we combine program search,
recognition model training, and DSL induction.
\begin{figure*}
\centering  \begin{tabular}{ll}
    \toprule
    Example programs in frontiers&Proposed subexpression\\\midrule
    \begin{tabular}{l}
      \code{(lambda (a b) (foldr b (cons "," a)}\\
      \phantom{\code{(lambda }}\code{(lambda (x z) (cons x z))))}\\
      \code{(lambda (a b) (foldr a b (lambda (x z) (cons x z))))}
    \end{tabular}&
    \begin{tabular}{l}
      \code{(foldr a b (lambda (x z) }\\
      \phantom{\code{(foldr a b }}\code{(cons x z)))}
      \end{tabular}
\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm proposes subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (left column), and can introduce new variables (right column: \code{a} and \code{b}). Here, the proposed subexpression appends two lists.} 
  \label{fragmentExample}
\end{figure*}

@@

\section{Sequence manipulating programs}\label{sequences}
We apply \system to list processing (Section~\ref{textSection}) and text editing (Section~\ref{listSection}).
For both these domains we use a bidirectional GRU~\cite{cho2014learning} for the recognition model,
and initially provide the system with a generic
set of list processing primitives: \code{foldr}, \code{unfold}, \code{if}, \code{map}, \code{length}, \code{index}, \code{=}, \code{+}, \code{-}, \code{0}, \code{1}, \code{cons}, \code{car}, \code{cdr}, \code{nil}, and \code{is-nil}.

@@

A complete repertoire of higher-order functions is a staple of functional programming standard libraries.
Although we provided \system with some of the standard higher-order functions,
like \code{foldr} and \code{unfold},
we did not include others.
When trained on these list functions,
our system rediscovers
and then reuses the higher-order function \code{filter} \textbf{Lucas: not sure if it actually does this! It would be cool if we can write something like this though.}

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b (lambda (x y) (cons x y))))))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %(lambda (lambda (fold $0 $0 (lambda (lambda (if (char-eq? $1 $3) empty (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some string editing learned subroutines }\label{textPrimitives}
  \end{table}
